Title

### 2013-03-22 inc/hardware/magnetic_lenses/maglens.tex
added polepiece-design drawing, this included moving some labels on the original schematic. Note that the figure description still needs to be changed!

Discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
These materials, called ``magnetic shielding'' since it captures the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
The field lines exit the shielding in the gap and it is this field that acts as the lens.
This gap can be much smaller than the total width of the coil turns, and yet contain in a field commensurate with the full coil.

\begin{figure}
  \centering
  \input{polepiece-schematic}
  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
    Side-view schematic of a shielded magnetic lens with a pole-piece.
    The windings (current loops) are contained within the shielding.
    The magnetic field created by these windings is contained inside the shielding except in the gap.
  }
\end{figure}

El-Kareh and El-Kareh \cite{el-kareh_electron_1970} provide an analysis of symmetric lenses with magnetic shielding and a pole-piece in chapter 8, with the most useful information being in section 8.7.
They define
\begin{subequations}
\begin{gather}
  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2 \\
  \beta \equiv k^2 V_r / (NI)^2 \label{eq:elkareh_k2}
\end{gather}
\end{subequations}
where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
\begin{equation}
  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V )
\end{equation}
They then derive a number of formulae and tabulate the results for a variety of parameters.
Using these tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.

%TODO check S and D as I seem to remember we have S=0.1in
%TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
For the UEM, I chose a bore diameter $D=2R=$0.1in and a pole-piece gap $S$=0.5in, for an ratio $S/D=0.2$.
By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

@@

>>>>

Discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
These materials, called ``magnetic shielding'' since it captures the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
The field lines exit the shielding in the gap and it is this field that acts as the lens.
This gap can be much smaller than the total width of the coil turns, and yet contain in a field commensurate with the full coil.
This geometry is presented schematically as \ref{fig:polepiece-schematic}.

\begin{figure}
  \centering
  \input{polepiece-schematic}
  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
    Side-view schematic of a shielded magnetic lens with a pole-piece.
    The windings (current loops) are contained within the shielding.
    The magnetic field created by these windings is contained inside the shielding except in the gap.
  }
  \label{fig:polepiece-schematic}
\end{figure}

El-Kareh and El-Kareh provide an analysis of symmetric lenses with magnetic shielding and a pole-piece (see Ref. \cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
They define
\begin{subequations}
\begin{gather}
  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2 \\
  \beta \equiv k^2 V_r / (NI)^2 \label{eq:elkareh_k2}
\end{gather}
\end{subequations}
where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
\begin{equation}
  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V )
\end{equation}
They then derive a number of formulae and tabulate the results for a variety of parameters.
Using these tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.

%TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
For the UEM, I chose a bore diameter $D=2R=$0.5in and a pole-piece gap $S$=0.1in, for an ratio $S/D=0.2$.
By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

@@

\begin{figure}
  \centering
  \input{polepiece-design}
  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
    Side-view schematic of a shielded magnetic lens with a pole-piece.
    The windings (current loops) are contained within the shielding.
    The magnetic field created by these windings is contained inside the shielding except in the gap.
  }
  \label{fig:polepiece-design}
\end{figure}
